\vspace{-1cm} \bfseries\large Phosphatidylinositol 4,5-bisphosphate regulates cilium transition zone maturation in \textit{Drosophila melanogaster}

### 2018-04-07 latex/paper.tex
Minor edits from me

    Ciliogenesis begins with assembly of a nascent TZ at the tip of
    the basal body (BB) \citep{reiter2012base}.
    During TZ maturation, its structure and protein
    constituents change, allowing for
    establishment of a compartmentalized space, bounded by the ciliary membrane
    and the TZ, where assembly of the axoneme, a microtubule-based structure
    that forms the ciliary core,
    and signalling can occur.
    In Drosophila, nascent TZs first assemble on BBs
    during early G2 phase in primary spermatocytes \citep{riparbelli2012assembly}.
    This occurs concomitantly with anchoring of cilia to the plasma membrane (PM),
    microtubule remodelling within the TZ \citep{gottardo2013cilium},
    and establishment of a
    ciliary membrane that will persist through meiosis
    \citep{riparbelli2012assembly} (Figure 1A).
    TZ maturation has been described in
    Paramecium \citep{aubusson2015transition},
    \textit{Caenorhabditis elegans} \citep{serwas2017centrioles} and
    Drosophila \citep{gottardo2013cilium},
    and is most readily observed by an increase in TZ length
    in the Drosophila male germline.

@@

    Vertebrate type I PIP kinase PIPKI$\gamma$ is important for cilium formation in cultured cells
    \citep{xu2016phosphatidylinositol}.
    The two Drosophila PIPKIs, Sktl and PIP5K59B,
    arose from recent duplication of
    the ancestral PIPKI gene,
    and are not orthologous to specific vertebrate PIPKI isoforms (Figure 2D).
    Sktl has diverged more than its paralog PIP5K59B and seems
    to be functionally related to PIPKI$\gamma$
    and the \textit{C. elegans} PPK-1 in having roles at cilia \citep{xu2014pipkigamma}.
    However, unlike the human PIPKI$\gamma$, which licenses TZ assembly
    by promoting CP110 removal from BBs \citep{xu2016phosphatidylinositol},
    our results suggest that Sktl functions in
    regulating TZ length but not TZ assembly. 
    Notably, neither inactivation nor overexpression of \textit{cp110} affects
    cilium formation in Drosophila, and Cp110
    is removed from BBs in early primary spermatocytes
    \citep{franz2013cp110}.

    % 
    % Results - 3
    % 
    \subsection*{Hyperelongated transition zones exhibit signs of functional defects}
    We next sought to examine whether TZ hyperelongation due to SigD expression
    affected TZ function.
    Following meiosis in the Drosophila male germline, TZs detach from
    the BB and migrate along the growing axoneme, maintaining a ciliary compartment
    at the distal-most {\raise.17ex\hbox{$\scriptstyle\sim$}}5$\mu$m where tubulin
    is incorporated into the axoneme
    \citep{basiri2014migrating, fabian2012drosophila}.
    As shown by Unc and Cep290 localization,
    TZs in \sigd{} were frequently incapable of detaching from
    BBs and migrating along axonemes despite
    axoneme and cell elongation
    (Figures 1E, 3A and 3B).
    Indeed, the previously reported ``comet-shaped'' Unc-GFP localization in \sigd{}
    \citep{wei2008depletion}
    persists during cell elongation after meiosis
    (Figure 3A, bottommost panel)
    despite elongation of the axoneme (Figure 1E).

    In Drosophila and humans, BBs consist of microtubule triplets
    \citep{jana2016drosophila, lattao2017centrioles},
    whereas axonemes contain microtubule doublets due to obstruction of
    C-tubules at the TZ \citep{gottardo2013cilium}.
    Consistent with a defect in this barrier and the presence of
    microtubule triplets in axonemes in \sigd{} \citep{wei2008depletion},
    a subset of cilia (<5\%) in \sigd{}
    contained puncta of Ana1 at the distal
    tips of TZs (Figure 3C).
    Treatment of germ cells with the microtubule-stabilizing drug
    Taxol increased the penetrance of this phenotype from <5\% in untreated cells
    to \textgreater25\% in cells treated with 4 $\mu$M Taxol (arrowheads in Figure 3D)
    without significantly affecting Cep290 length (Figure 3E).
    Taxol-treated controls did not exhibit TZ-distal Ana1 puncta ($p$ <0.01 at 5\% penetrance).
    Fluorescently-tagged Asterless (CEP152 homolog), a pericentriolar protein
    \citep{blachon2008drosophila, dzhindzhev2010asterless},
    did not localize to TZ-distal puncta in \sigd{} ($p$ <0.01)
    suggesting that these TZ-distal
    sites are not fully centriolar in protein composition.
    Taxol has been hypothesized to disrupt TZ maturation
    by inhibiting microtubule remodelling in the Drosophila male germline \citep{riparbelli2012assembly, riparbelli2013unique}.
    Similar to \sigd{},
    Taxol-treated male germ cells assemble extremely long axonemes
    that contain triplet microtubules
    \citep{riparbelli2013unique}, further supporting a functional relationship between
    \PIP{} and microtubule reorganization in TZ maturation.

    % 
    % Results - 4
    % 
    \subsection*{The onion rings (\textit{onr}) mutant decouples defects found in cells with reduced levels of \PIP{}}
    Male flies homozygous for the \textit{onion rings}
    (\textit{onr}) mutant of Drosophila Exo84 are sterile and exhibit defects
    in cell elongation and polarity
    similar to \sigd{} \citep{wei2008depletion}.
    Exo84 is a component of the octameric exocyst complex, which binds
    \PIP{} and regulates membrane trafficking at the PM 
    \citep{he2007exo70}.
    To investigate whether defects in TZ hyperelongation could be explained by
    defective Exo84 function, we examined TZs in \textit{onr} mutants.
    Unlike \sigd{}, \textit{onr} did not display hyperelongated TZs (Figure 4A),
    suggesting that Exo84 is dispensable for TZ maturation.

@@

    The majority of \PIP{} at the PM is produced by PIPKIs
    \citep{balla2013phosphoinositides, hammond2012pi4p}.
    In this study, we showed that mutation of the PIPKI Sktl
    induced hyperelongated TZs
    and that expression of Sktl could suppress TZ hyperelongation in \sigd{},
    with some cells showing cilium-autonomous suppression,
    suggesting Skittles might function \textit{in situ} to regulate TZ length.
    In humans, \textit{PIPKI$\gamma$} is linked to lethal congenital contractural
    syndrome type 3 (LCCS3), which has been suggested to represent a ciliopathy
    \citep{xu2016phosphatidylinositol}.
    The recent discovery of a role for LCCS1-associated GLE1 protein in cilium
    function \citep{jao2017role} corroborates this hypothesis.
    Our data support the idea that PIPKIs might represent ciliopathy-associated genes
    or genetic modifiers of disease.

@@

    \subsection*{Statistical methods}
    Statistical analysis and graphing was performed using R (version 3.4)
    \citep{r}.
    A Gaussian jitter was applied when plotting
    results in Figures 1 and 2 for clearer visualization of trends,
    but raw data was used for all analyses.
    Statistical tests for ``absence of phenotype'' were computed using a binomial test
    under the assumption that the probability of the phenotype occuring was fixed.
    All $t$-tests were unpaired and two-sided with Welch's correction for unequal variances.
    $n$ represents the pooled number of samples (individual cilia) from multiple flies.
    A significance level of 0.01 was fixed \textit{a priori} for all classical analyses.
    All raw data and code for analysis and plotting can be found online
    at \url{http://www.github.com/alindgupta/germline-paper/}.

@@

    \section*{Acknowledgements}
    We are grateful to Brian Ciruna for insightful discussions,
    B{\'e}n{\'e}dicte Durand, Tomer Avidor-Reiss, Antoine Guichet and Maurice Kernan for providing
    unpublished fly stocks, and to
    B{\'e}n{\'e}dicte Durand and Bill Trimble
    for comments on the manuscript.
    This research was supported by a grant from the Canadian Institutes for
    Health and Research to J.A.B (MOP \#130437).
    A.G. was supported by a
    University of Toronto Open Fellowship and
    an Ontario Graduate Scholarship.

    \section*{Author contributions}
    A.G. and J.A.B. conceived the project.
    A.G. performed all the experiments and analysis, and wrote the manuscript.
    L.F. generated GFP-Exo70 flies.
    J.A.B. edited the manuscript and supervised the project.

@@

    \item Phylogenetic tree of PIP5Ks showing conservation of cilium-associated functions.

@@

    \item GFP-tagged Unc is unable to detach from the basal body and migrate in spermatids expressing \sigd{} (white arrowhead).

@@

    \textbf{\itshape Figure 4. The \textit{onion rings} (\textit{onr}) allele of Exo84 decouples TZ hyperlongation from loss of plasma membrane contacts.}

@@

    \item Cells mutant for \textit{onr} do not display hyperelongated acetylated tubulin signal at the cilium (white arrowhead).

>>>>

    Ciliogenesis begins with assembly of a nascent TZ at the tip of
    the basal body (BB) \citep{reiter2012base}.
    During TZ maturation, its structure and protein
    constituents change, allowing for
    establishment of a compartmentalized space, bounded by the ciliary membrane
    and the TZ, where assembly of the axoneme, a microtubule-based structure
    that forms the ciliary core,
    and signalling can occur.
    In Drosophila, nascent TZs first assemble on BBs
    during early G2 phase in primary spermatocytes \citep{riparbelli2012assembly}.
    This occurs concomitantly with anchoring of cilia to the plasma membrane (PM),
    microtubule remodelling within the TZ \citep{riparbelli2013unique, gottardo2013cilium},
    and establishment of a
    ciliary membrane that will persist through meiosis
    \citep{riparbelli2012assembly} (Figure 1A).
    TZ maturation has been described in
    Paramecium \citep{aubusson2015transition},
    \textit{Caenorhabditis elegans} \citep{serwas2017centrioles} and
    Drosophila \citep{gottardo2013cilium},
    and is most readily observed by an increase in TZ length
    in the Drosophila male germline.

@@

    Vertebrate type I PIP kinase PIPKI$\gamma$ has previously been shown to be important for cilium formation in cultured cells
    \citep{xu2016phosphatidylinositol}.
    The two Drosophila PIPKIs, Sktl and PIP5K59B,
    arose from recent duplication of
    the ancestral PIPKI gene,
    and are not orthologous to specific vertebrate PIPKI isoforms (Figure 2D).
    Sktl has diverged more than its paralog PIP5K59B and seems
    to be functionally related to PIPKI$\gamma$
    and the \textit{C. elegans} PPK-1 in having roles at cilia \citep{xu2014pipkigamma}.
    However, unlike the human PIPKI$\gamma$, which licenses TZ assembly
    by promoting CP110 removal from BBs \citep{xu2016phosphatidylinositol},
    our results suggest that Sktl functions in
    regulating TZ length but not TZ assembly. 
    Notably, neither inactivation nor overexpression of \textit{cp110} affects
    cilium formation in Drosophila, and Cp110
    is removed from BBs in early primary spermatocytes
    \citep{franz2013cp110}.

    % 
    % Results - 3
    % 
    \subsection*{Hyperelongated transition zones exhibit functional defects}
    We next sought to examine whether TZ hyperelongation due to SigD expression
    affected TZ function.
    Following meiosis in the Drosophila male germline, TZs detach from
    the BB and migrate along the growing axoneme, maintaining a ciliary compartment
    at the distal-most {\raise.17ex\hbox{$\scriptstyle\sim$}}5$\mu$m where tubulin
    is incorporated into the axoneme
    \citep{basiri2014migrating, fabian2012drosophila}.
    As shown by Unc and Cep290 localization,
    TZs in \sigd{} were frequently incapable of detaching from
    BBs and migrating along axonemes despite
    axoneme and cell elongation
    (Figures 1E, 3A and 3B).
    Indeed, the previously reported ``comet-shaped'' Unc-GFP localization in \sigd{}
    \citep{wei2008depletion}
    persists during cell elongation after meiosis
    (Figure 3A, bottommost panel)
    despite elongation of the axoneme (Figure 1E).

    In Drosophila and humans, BBs consist of microtubule triplets
    \citep{jana2016drosophila, lattao2017centrioles},
    whereas axonemes contain microtubule doublets due to obstruction of
    C-tubules at the TZ \citep{gottardo2013cilium}.
    Consistent with a defect in this barrier and the presence of
    microtubule triplets in axonemes in \sigd{} \citep{wei2008depletion},
    a subset of cilia (<5\%) in \sigd{}
    contained puncta of Ana1 at the distal
    tips of TZs (Figure 3C).
    Treatment of germ cells with the microtubule-stabilizing drug
    Taxol increased the penetrance of this phenotype from <5\% in untreated cells
    to \textgreater25\% in cells treated with 4 $\mu$M Taxol (arrowheads in Figure 3D)
    without significantly affecting Cep290 length (Figure 3E).
    Taxol-treated controls did not exhibit TZ-distal Ana1 puncta ($p$ <0.01 at 5\% penetrance).
    Fluorescently-tagged Asterless (CEP152 homolog), a pericentriolar protein
    \citep{blachon2008drosophila, dzhindzhev2010asterless},
    did not localize to TZ-distal puncta in \sigd{} ($p$ <0.01)
    suggesting that these TZ-distal
    sites are not fully centriolar in protein composition.
    Taxol has been hypothesized to disrupt TZ maturation
    by inhibiting microtubule remodelling in the Drosophila male germline \citep{riparbelli2013unique}.
    Similar to \sigd{},
    Taxol-treated male germ cells assemble extremely long axonemes
    that contain triplet microtubules
    \citep{riparbelli2013unique}, further supporting a functional relationship between
    \PIP{} and microtubule reorganization in TZ maturation.

    % 
    % Results - 4
    % 
    \subsection*{The onion rings (\textit{onr}) mutant decouples defects found in cells with reduced levels of \PIP{}}
    Male flies homozygous for the \textit{onion rings}
    (\textit{onr}) mutant of Drosophila \textit{exo84} are sterile and exhibit defects
    in cell elongation and polarity
    similar to \sigd{} \citep{wei2008depletion}.
    Exo84 is a component of the octameric exocyst complex, which binds
    \PIP{} and regulates membrane trafficking at the PM 
    \citep{he2007exo70}.
    To investigate whether defects in TZ hyperelongation could be explained by
    defective Exo84 function, we examined TZs in \textit{onr} mutants.
    Unlike \sigd{}, \textit{onr} did not display hyperelongated TZs (Figure 4A),
    suggesting that Exo84 is dispensable for TZ maturation.

@@

    The majority of \PIP{} at the PM is produced by PIPKIs
    \citep{balla2013phosphoinositides, hammond2012pi4p}.
    In this study, we showed that mutation of the PIPKI Sktl
    induced hyperelongated TZs
    and that expression of Sktl could suppress TZ hyperelongation in \sigd{},
    with some cells showing cilium-autonomous suppression,
    suggesting Sktl might function \textit{in situ} to regulate TZ length.
    In humans, \textit{PIPKI$\gamma$} is linked to lethal congenital contractural
    syndrome type 3 (LCCS3), which has been suggested to represent a ciliopathy
    \citep{xu2016phosphatidylinositol}.
    The recent discovery of a role for LCCS1-associated GLE1 protein in cilium
    function \citep{jao2017role} corroborates this hypothesis.
    Our data support the idea that PIPKIs might represent ciliopathy-associated genes
    or genetic modifiers of disease.

@@

    \subsection*{Statistical methods}
    Statistical analysis and graphing was performed using R (version 3.4).
    A Gaussian jitter was applied when plotting
    results in Figures 1 and 2 for clearer visualization of trends,
    but raw data was used for all analyses.
    Statistical tests for ``absence of phenotype'' were computed using a binomial test
    under the assumption that the probability of the phenotype occuring was fixed.
    All $t$-tests were unpaired and two-sided with Welch's correction for unequal variances.
    $n$ represents the pooled number of samples (individual cilia) from multiple flies.
    A significance level of 0.01 was fixed \textit{a priori} for all classical analyses.
    All raw data and code for analysis and plotting can be found online
    at \url{http://www.github.com/alindgupta/germline-paper/}.

@@

    \section*{Acknowledgements}
    We are grateful to Brian Ciruna for insightful discussions,
    to B{\'e}n{\'e}dicte Durand, Tomer Avidor-Reiss, Antoine Guichet and Maurice Kernan for providing
    fly stocks, and to
    B{\'e}n{\'e}dicte Durand and Bill Trimble
    for critical comments on the manuscript.
    This research was supported by a grant from the Canadian Institutes for
    Health and Research to J.A.B (MOP \#130437).
    A.G. was supported by a
    University of Toronto Open Fellowship and
    an Ontario Graduate Scholarship.

    \section*{Author contributions}
    A.G. and J.A.B. conceived the project.
    A.G. performed all the experiments and analyses, and wrote the manuscript.
    L.F. generated GFP-Exo70 flies.
    J.A.B. edited the manuscript and supervised the project.

@@

    \item Phylogenetic tree of PIP5Ks showing evolutionary conservation of cilium-associated functions.

@@

    \item Unc-GFP is unable to detach from the basal body and migrate in spermatids expressing \sigd{} (white arrowhead).

@@

    \textbf{\itshape Figure 4. The \textit{onion rings} (\textit{onr}) allele of \textit{exo84} decouples TZ hyperlongation from loss of plasma membrane contacts.}

@@

    \item \textit{onr} mutant cells do not display hyperelongated acetylated tubulin signal at the cilium (white arrowhead).
